### 2021-06-06 paper/parlr_paper_1.Rmd
first draft of paper

---
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
  html_document:
    df_print: paged
header-includes: \usepackage{hyperref}
linestretch: 1
link-citations: yes
linkcolor: blue
fontfamily: mathpazo
fontsize: 12pt
---
\begin{flushright} 
	\end{flushright}
	\begin{center} \textbf{Parlr: Parallelized Record Linkage in R}

	Brian Kundinger

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

@@

Record linkage is the task of identifying duplicate records across multiple data sources. This task is straightforward when provided unique identifiers or highly reliable information, but becomes difficult when data sources are plagued with error. Additionally, many record linkage tasks are large in nature, and many existing methods are too computational intensive to be used in practice. In this paper, we propose a Bayesian model for record linkage that is amenable to parallel computing, hashing techniques and efficient storage mechanisms, that is scalable to large linkage tasks.

The proposed method is inspired by the "Beta Record Linkage" method proposed by Mauricio Sadinle's 2017 "Bayesian Estimation of Bipartite Matching for Record Linkage". As such, we demonstrate our method through a comparison of our methods against Sadinle's on the simulation study he designed and the El Salvador homocide case study he utilized. We include one additional simulation study to demonstrate the power of our method on large datasets. In what follows, Section 2 describes previous work in record linkage that lays the foundation for our method, Section 3 provides our model and the Gibbs sampler used for posterior inference, and Section 4 describes strategies used for more efficient and scalable implementation. Then, Sections 5 and 6 demonstrate the method through simulations and a case study. Lastly, our discussion highlights open questions in the record linkage liturature and motivates future work.

@@

Most record linkage techniques are derived from the seminal 1969 paper by Fellegi and Sunter: "A Theory for Record Linkage". The defining characteristic of their model was to transform the sets of records, which often contain text data that is difficult to model, into sets of comparison vectors governed by parameters that can be more easily estimated. Concretely, if files $A$ and $B$ have $n_A$ and $n_B$ records respectively, and if the files share $F$ fields in common upon which to base the linkage, the Fellegi and Sunter approach generates a $n_A n_B \times F$ matrix $\Gamma$, which contains similarity scores between each pair of records across datasets. We say $\gamma_{ij}$ is the comparison vector for record $i \in A$ and record $j \in B$, with $\gamma_{ij}^f \in \{1, \ldots, L_f\}$ providing their similarity score on the $f^{th}$ field. For ease of modeling and computation, we restrict these similarity scores to be discrete, ordinal variables, and the construction of these is left to the modeler. We adopt the convention that $\gamma_{ij}^f = 1$ corresponds to the highest level of agreement and $\gamma_{ij}^f = L_f$ corresponds to the lowest. It is common to use use binary 1-2 variables to indicate exact matching, and 1-2-3 variables to provide an option for partial matching. For text data, we calculate similarity based on Levenstein distance or some other text similarity score, and bin these scores to integers for use in the model.

The likelihood in the original Fellegi Sunter model is a mixture model through which each record pair is independently classified as a match or nonmatch. This independent classification often leads to sets of matches that break expectations of transisitivity, and Jaro devised an procedure to reduce a set of conflicting matches to the mostly likely set of one-to-one matchings. One key strength of Sadinle's `BRL` is that it is garunteed to produce a set of bipartite matchings that respect one-to-one assumptions--without the need for any post-hoc procedures. Specifically, in each iteration of his Gibbs sampler, he considers each record $j\in B$, removes from consideration the records $i\in A$ that have already been matched, and then samples a potential link. However, accounting for these dependencies throughout the linkage process is computationally burdensome, leaving `BRL` only suitable for small to moderate linkage problems.

<!-- Two modern adaptations of the Fellegi Sunter model are important for understanding `parlr` and its contribution. In 2019, Enamorado et al proposed `fastlink`, a method that closely followed the modelling assumptions of Fellegi and Sunter, but used innovating hashing techniques to greatly enhance its speed and scale. Importantly, they maintained Fellegi and Sunter's independent matching assumption, in which the matching status of any $(i,j)$ pair is made independently of the matching of any other $(i', j')$. This leads to many matchings that violate "one-to-one" considerations that are often implicit in the data, which need to be resolved in a post-processing step. In contrast, Sadinle (2017) proposed "Beta Record Linkage for Bipartite Matching" and the accompanying `BRL` package which strictly enforces one-to-one matching. Specifically, in each iteration of his Gibbs sampler, he considers each record $j\in B$, removes from consideration the records $i\in A$ that have already been matched, then samples a potential link. However, accounting for these dependencies throughout the linkage process is computationally burdensome, leaving `BRL` only suitable for small to moderate linkage problems.  -->

@@

Our notation and assumptions closely follow those of Sadinle. Denote two files as $A$ and $B$, with $n_A$ and $n_B$ records respectively, and with records indexed as $i \in \{1, \ldots, n_A\}$ in $A$ and $j \in \{1, \ldots, n_B\}$ in $B$. Without loss of generality, label the files such that $n_A \geq n_B$. We also assume there are no duplicates within files, only across. For each record pair under consideration, we generate a comparison vector $\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}$, where  $F$ is the number of fields used in the linkage and each $\gamma_{ij}^f$ takes on a value $l \in \{1, \ldots, L_f\}$ indicating the level agreement between the two records on a specified field.

@@

$$\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda}) $$
The prior for $Z_j$ has equal probability of matching to all records $i\in A$, and non-matching probability governed by $\lambda$. Therefore a  $\lambda \sim \text{Beta}(1, 1)$ corresponds to a prior belief that nonmatches and matches are equally likely, and a $\lambda \sim \text{Beta}(1, \frac{1}{n_A})$ prior corresponds to a uniform prior on the labeling of $\mathbf{Z}$.

Here the reader should the relationship between our proposed model and that of Sadinle 2017. In his model, Sadinle constructs a prior distribution on the entire $\mathbf{Z}$ vector, which induces a Gibbs sampler that strictly enforces one-to-one matching. In particular, this sampler removes previously matches records from the set of candidate records when sampling $Z_j$, creating a dependency that makes the sampler *inherently serial*. We however use independent priors for each $Z_j$, thereby weakening the one-to-one requirement; our sampler does ensure that each record in $B$ can be matched to at most one record in $A$, but allows for the possiblity that multiple records in $B$ match to the same record in $A$. We explore the ramifications of this distinction with the El Salvador case study. With this change, components of the $\mathbf{Z}$ vector can be computed in parallel, allowing for significant computational gains.  More importantly, since only the agreement pattern of $Z_j$ is used for calculations within the Gibbs sampler, and not the particular record label, we can conduct this sampling only at the level of the unique agreement patterns, offering even more computational savings. This will be explained further in \textcolor{red}{Section XX}.

@@

We calculate a Bayes estimate $\hat{\mathbf{Z}}$ for the linkage parameter $\mathbf{Z}$ by assigning different positive losses to different types of errors, and minimizing posterior loss. Specifically, we minimize the quantity $L(\hat{\mathbf{Z}}, \mathbf{Z}) = \sum_{j=1}^{n_B} L(\hat{Z_j}, Z_j)$ where

$$L(\hat{Z_j}, Z_j)=\begin{cases} 
    0  & \text{if } Z_j = \hat{Z_j}; \\
    n_A + 1,  & \text{if record } j\in B \text{ does not have a match in file } A; \\
\end{cases}$$

<!-- Our Gibbs sampler provides posterior samples of $\mathbf{Z}$ which we use to make our final decisions about the linkage structure. In the case where we false matches and missed matches contribute the same loss, we declare $(i,j)$ to be a match whenever $P(Z_j = i) > \frac{1}{2}$ according to these posterior samples. We can create more elaborate decisions by attributing different loss values to different kinds of error, and allowing pairings with middling posterior probabilities to be left unlabeled by the algorithm so that the modeler can address those manually. For further discussion, see Sadinle 2017. -->

(Adapt from Sadinle)

Since our Gibbs procedure does not strictly enforce one-to-one matching, it is possible for the final Bayes estimate to link multiple records in $B$ to one record in $A$. The modeler can either report both such matches (with their respective posterior match probabilities), or resolve these conflicts by accepting only the match with highest posterior probability. Rigorous proof of this resolution procedure is deferred to future work.

@@

Broadly speaking, we increase our computational efficiency by recognizing that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector. Let $\mathcal{H}$ be the set of unique agreement patterns in the data, let $P$ denote the total number of unique agreement patterns. Note that $P$ is bounded above by $\prod_{f=1}^F L_f$, and that this bound does not scale with $n_A$ or $n_B$. Prior to processing the data, we identify all $P$ patterns in $\mathcal{H}$ and enumerate them $h_1, \ldots, h_P$. When the $(i,j)$ pair exhibits the $p^{th}$ agreement pattern, we say $(i,j) \in h_p$. Wherever possible, we conduct calculations over these $P$ agreement patterns rather than the $n_A \times n_B$ record pairs.

@@

The classic Fellegi Sunter method represents the $\gamma_{ij}$ comparison vector as vector of length $F$, with each component $\gamma_{ij}^f$ taking on values in  $\{0, \ldots, L_f - 1\}$. To ease computations, we instead represent the comparison as a concatenation of $F$ many binary indicator vectors of lengths $L_f$. For example, if $L_1 = L_2 = 2$ and $L_3 = 3$, then $\gamma_{ij} = (1, 0, 2)$ under the classical framework becomes $\gamma_{ij} = (0, 1, 1, 0, 0, 0, 1)$ under our framework. This is a bijective transformation that does not change the meaning of the data, but this representation will ease calculations and posterior updates.

In the classic Fellegi Sunter framework, $\Gamma$ is a $n_A n_B \times F$ matrix, with each row providing the comparison vector for a different $(i,j)$ pair. We however do not store these comparison vectors themselves, but instead only a hashed value corresponding to the agreement pattern of the $(i, j)$ pair. We store this information in a nested list $\tilde{\Gamma}$ where the $p^{th}$ component of the $j^{th}$ list contains a vector of records in $A$ that share agreement pattern $p$ with record $j \in B$. For each $p$, we also calculate $H_p = \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} \mathbf{1}_{(i,j) \in h_p}$, the total instances of agreement pattern $p$ throughout the data, and also for each $j$, we  calculate $H_{p_j} = \sum_{i=1}^{n_A} \mathbf{1}_{{(i,j) \in h_p}}$ the instances of agreement pattern $p$ among the comparison vectors between record $j \in B$ and each of the $n_A$ records in $A$.

For large data, we can partition the two datasets $A$ and $B$ into smaller blocks $\{A_m\}$ and $\{B_m\}$ for more manageable computations. On a single machine, we can read-in data sequentially, conduct hashing, collect results, and delete the original data from memory before continuing with the next chunk of data. With multiple cores or multiple machines, this can be done in parallel. Storing this hashed information still becomes burdensome for large data, but this hashing method greatly expands the capabilities of the Fellegi-Sunter framework.

@@

To address this, we propose a new strategy called *storage efficient indexing* (SEI). In standard indexing, the modeler decides a certain criteria that they expect all true matching pairs to satisfy, and a priori label any record pairs that do not meet that criteria as non-matching. For example, one might only consider pairs with a certain similarity score on a field deemed to be important (like first name), or only pairs with exact matching on a specified number of fields. While generally chosen to be be quite loose, establishing these criteria requires knowledge of the problem and invites room for human error. We propose a method of reducing the comparison space (and reducing storage requirements) without these drawbacks. Note that all records of the same agreement pattern have the same probability when sampling $Z_j$. Therefore we know that records belonging to an $h_p$ such that $H_{p_j}$ is large are very unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate, even without considering the form of the agreement pattern itself. In random indexing, rather than store all of these unlikely record labels, we choose to store only a small number $R$ of them. Posterior calculations still attribute the appropriate weight to all records through the summary statistics $H_{p_j}, H_p^m$, and $H_p^u$. Rather than storing $n_A \times n_B$ record labels, SEI allows us to store at most $n_B \times P \times R$ labels, regardless of how large $n_A$ is. Coupled with the partitioning strategy for creating comparison vectors, this greatly expands the size of the linkage tasks we can undertake.

@@

<!-- Sadinle uses a prior for $\mathbf{Z}$ that induces the a full conditional for $Z_j$ that strictly enforces one-to-one matching, given by  -->

<!-- $$p\left(Z_j^{(s+1)}  = i| \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}}\right) \propto -->
<!-- \begin{cases}  -->
<!--     w_{ij} \mathbf{1}_{i \neq Z_{j'}; \; j' \neq j} & i \leq n_A; \\ -->
<!--      D_{-j} \frac{n_B - D_{-j} + \beta_{\lambda}}{D_{-j} + \alpha_{\lambda}} & i  = n_A + 1 \\ -->
<!-- \end{cases}$$ -->

<!-- The indicator removes previously matched records from the set of candidate records, and $D_{-j}:=\sum_{B \setminus j} $ updates iteratively after each matching decision, $n_B$ many times per Gibbs iteration. This process is *inherently serial* and becomes computationally burdensome for large data.  -->

<!-- Sadinle uses a prior for $\mathbf{Z}$ that induces the a full conditional for $Z_j$ that strictly enforces one-to-one matching. In particular, this sampler removes previously matches records from the set of candidate records when sampling $Z_j$, creating a dependency that makes the sampler *inherently serial*. By weakening the one-to-one requirement, our full conditional for $Z$ does not depend on the rest of the $\mathbf{Z}_{-j}$ vector, and thus can be computed in parallel. More importantly, since only the agreement pattern of $Z_j$ is used for calculations within the Gibbs sampler, and not the particular record label, we can conduct this sampling only at the level of the unique agreement patterns. This boosts computation time far greater than parallelization.  -->

\underline{Updating $\mathbf{Z}$:} Although sampling $Z_j$ from a the full conditional provided earlier is conceptually straightforward, it becomes computational burdensome when $n_A$ is larger. The reader can confirm that sampling a value from a large set of unequal probabilities becomes difficult in most programming languages. To speed up computation, we break this sampling step into two simpler steps. First, we calculate the Fellegi Sunter weight $w_{h_p}$ associated with each unique pattern and sample the agreement pattern between $j$ and its potential match. Second, we sample the record label uniformly among records associated with that agreement pattern for that particular $j\in B$. More concretely, define $h(Z_j)$ to be the agreement pattern between $j$ and its potential match, and say $h(Z_j) = h_{P+1}$ when $Z_j = n_A + 1$. Then,

$$h\left(Z_j^{(s+1)}\right) | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
\begin{cases} 
    w_{h_p}\times H_{p_j}  & p \leq P; \\
     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &   p = P + 1 \\
\end{cases}$$
<!-- $$p\left(Z_j^{(s+1)} = i\right) = \frac{1}{\left|H\left(Z_j^{(s+1)}\right)\right|}; \;\;i \text{ such that } (i, j) \in H\left(Z_j^{(s+1)}\right)$$ -->

Lastly, we recognize that all posterior updates are governed by the agreement patterns of the record pairs rather than the record labels themselves. Thuse We complete the entire Gibbs procedure at the level of the $P$ agreement patterns. After, we can back-fill the records corresponding to the agreement patterns by sampling uniformly at random among candidate records stored in $\tilde{\Gamma}$. Sampling uniformly is computational simple even for large sets of candidate records, but this step can also be parallelized when working with large data.

@@

Comparing our method against `BRL` on these same simulated datasets, we find that our method only has weakened performance in the most extreme scenario of very high errors and very high overlap across files. We use the following measures of accuracy in our analysis: although using only recall and precision is common in the literature, we find it useful to have a single metric the gauge overall performance, and thus also include F measure.

@@

```{r fig.align='center'}
knitr::include_graphics("../notes/figures/sadinle_sim_plot.png")
```

@@

To demonstrate speed, we generate comparison vectors from pre-specified distributions so that we can easily increase the size of the linkage problem. Distributions are meant to emulate the behavior of similarity scores across first name, last name, and day, month, and and year of birth, with 1 indicating an exact match on a particular field and 0 indicating a nonmatch. We simulate these data for different values of $n_A$ and $n_B$, and compare the run-time of `parlr` against `BRL`. Note that the number of unique patterns $P$ is bounded above by $2^5 = 32$, a bound which is consistently attained in the larger simulations.

We see that at low data size, `BRL` outperforms, but that `parlr` is significantly faster at handling larger data. In particular, run-time for `BRL` seems to grow quadratically (or linearly with the size of both $A$ and $B$) while run-time for `parlr` seems to grow linearly (in the size of $B$).

\begin{table}[ht]
\centering
\begin{tabular}{rrr}
  \hline
 & m & u \\ 
  \hline
fname\_0 & 0.05 & 0.99 \\ 
  fname\_1 & 0.95 & 0.01 \\ 
  lname\_0 & 0.05 & 0.99 \\ 
  lname\_1 & 0.95 & 0.01 \\ 
  day\_0 & 0.05 & 0.97 \\ 
  day\_1 & 0.95 & 0.03 \\ 
  month\_0 & 0.05 & 0.92 \\ 
  month\_1 & 0.95 & 0.08 \\ 
  year\_0 & 0.05 & 0.93 \\ 
  year\_1 & 0.95 & 0.07 \\ 
   \hline
\end{tabular}
\end{table}

```{r fig.align='center'}
knitr::include_graphics("../notes/figures/sadinle_speed_plot.png")
```

The above discussion suggests that for fixed $n_B$, computation time should remain mostly constant with growing $n_A$. Simulation study suggests that this is true. In the plot below, fixing $n_B = 500$, we see linear growth for the run-time under `BRL` as $n_A$ increases, with much more static run-time under `parlr`. The slight increases in run-time that we do see are due primarily to the hashing step, which again can be run in parallel for large data.

```{r fig.align='center'}
knitr::include_graphics("../notes/figures/speed_plot_fixed_nB.png")
```

>>>>

---
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
  html_document:
    df_print: paged
header-includes: \usepackage{hyperref}
linestretch: 1
link-citations: yes
linkcolor: blue
fontfamily: mathpazo
fontsize: 12pt
---
\begin{flushright} 
	\end{flushright}
	\begin{center} \textbf{Efficient and Scalable Entity Resolution with Fast Beta Linkage  (fabl)}

	Brian Kundinger, Rebecca Steorts, and Jerome Reiter

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align='center')
```

@@

Entity resolution (or record linkage) is the task of identifying duplicate records across multiple data sources. This task is straightforward when provided unique identifiers or highly reliable information, but becomes difficult when data sources are plagued with error. Additionally, many entity resolution tasks are large in nature, and many existing methods are too computational intensive to be used in practice. In this paper, we propose a Bayesian model for record linkage that uses hashing techniques and efficient storage mechanisms, and is amenable to parallelization, making it scalable to large linkage tasks.

The proposed method is inspired by the *Beta Record Linkage* (`BRL`) method proposed in Mauricio Sadinle's 2017 "Bayesian Estimation of Bipartite Matching for Record Linkage". To recognize this lineage, we name our method *Fast Beta Linkage* (`fabl`, pronounced "fable"). In what follows, Section 2 describes previous work in record linkage that lays the foundation for our method, Section 3 provides our model and the Gibbs sampler used for posterior inference, and Section 4 describes strategies used for more efficient and scalable implementation. Then, Section 5  demonstrates the accuracy and speed of out method through two simulation studies, and Section 6 reveals interesting properties of the method through a case study of El Salvadoran casuality records. Lastly, our discussion highlights open questions in the record linkage liturature and motivates future work.

@@

Most record linkage techniques are derived from the seminal 1969 paper by Fellegi and Sunter: "A Theory for Record Linkage". The defining characteristic of their approach was to transform the sets of records, which often contain text data that is difficult to model, into sets of comparison vectors governed by parameters that can be more easily estimated. Concretely, if files $A$ and $B$ have $n_A$ and $n_B$ records respectively, and if the files share $F$ fields in common upon which to base the linkage, the Fellegi and Sunter approach generates an $n_A n_B \times F$ matrix $\Gamma$, which contains similarity scores between each pair of records across datasets. We say $\gamma_{ij}$ is the comparison vector for record $i \in A$ and record $j \in B$, with $\gamma_{ij}^f \in \{1, \ldots, L_f\}$ providing their similarity score on the $f^{th}$ field. For ease of modeling and computation, we restrict these similarity scores to be discrete, ordinal variables, and the construction of these is left to the modeler. We adopt the convention that $\gamma_{ij}^f = 1$ corresponds to the highest level of agreement and $\gamma_{ij}^f = L_f$ corresponds to the lowest. It is common to use use binary 1-2 variables to indicate exact matching, and 1-2-3 variables to provide an option for partial matching. For text data, we calculate similarity based on Levenstein distance or some other text similarity score, and bin these scores to integers for use in the model. Jaro (1989) provided guidelines for appropriate binning for the comparison vectors, but details of implementation are generally context specific.

The likelihood in the original Fellegi Sunter model is a mixture model through which each record pair is independently classified as a match or nonmatch. This independent classification often leads to sets of matches that break expectations of transitivity, which is undesirable. Although Jaro devised an procedure to reduce a set of conflicting matches to the mostly likely set of one-to-one matchings, the abundance of false matches can bias the estimation of other parameters of the model.

One key strength of Sadinle's `BRL` is that it is garunteed to produce a set of bipartite matchings that respect one-to-one assumptions--without the need for any post-hoc procedures. Specifically, in each iteration of his Gibbs sampler, he considers each record $j\in B$, removes from consideration the records $i\in A$ that have already been matched, and then samples a potential link. However, accounting for these dependencies throughout the linkage process is computationally burdensome, leaving `BRL` only suitable for small to moderate linkage problems.

<!-- Two modern adaptations of the Fellegi Sunter model are important for understanding `fabl` and its contribution. In 2019, Enamorado et al proposed `fastlink`, a method that closely followed the modelling assumptions of Fellegi and Sunter, but used innovating hashing techniques to greatly enhance its speed and scale. Importantly, they maintained Fellegi and Sunter's independent matching assumption, in which the matching status of any $(i,j)$ pair is made independently of the matching of any other $(i', j')$. This leads to many matchings that violate "one-to-one" considerations that are often implicit in the data, which need to be resolved in a post-processing step. In contrast, Sadinle (2017) proposed "Beta Record Linkage for Bipartite Matching" and the accompanying `BRL` package which strictly enforces one-to-one matching. Specifically, in each iteration of his Gibbs sampler, he considers each record $j\in B$, removes from consideration the records $i\in A$ that have already been matched, then samples a potential link. However, accounting for these dependencies throughout the linkage process is computationally burdensome, leaving `BRL` only suitable for small to moderate linkage problems.  -->

@@

Denote two files as $A$ and $B$, with $n_A$ and $n_B$ records respectively, and with records indexed as $i \in \{1, \ldots, n_A\}$ in $A$ and $j \in \{1, \ldots, n_B\}$ in $B$. Without loss of generality, label the files such that $n_A \geq n_B$. We assume there are no duplicates within files, only across. For each record pair under consideration, we generate a comparison vector $\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}$, where  $F$ is the number of fields used in the linkage and each $\gamma_{ij}^f$ takes on a value $l \in \{1, \ldots, L_f\}$ indicating the level agreement between the two records on a specified field.

@@

$$\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda}) $$
The prior for $Z_j$ has equal probability of matching to all records $i\in A$, and non-matching probability governed by $\lambda$. Therefore a  $\lambda \sim \text{Beta}(1, 1)$ corresponds to a prior belief that nonmatches and matches are equally likely, and a $\lambda \sim \text{Beta}\left(1, \frac{1}{n_A}\right)$ prior corresponds to a uniform prior on the labeling of $\mathbf{Z}$.

Here the reader should the relationship between our proposed model and that of Sadinle 2017. In his model, Sadinle constructs a prior distribution on the entire $\mathbf{Z}$ vector, which induces a Gibbs sampler that strictly enforces one-to-one matching. In particular, this sampler removes previously matches records from the set of candidate records when sampling $Z_j$, creating a dependency that makes the sampler *inherently serial*. We however use independent priors for each $Z_j$, creating a sampler that is *perfectly parallelizable*, allowing for significant computational gains. More importantly, since only the agreement pattern of $Z_j$ is used for calculations within the Gibbs sampler, and not the particular record label, we can conduct this sampling only at the level of the unique agreement patterns, offering even more computational savings. In doing so however, we thereby weaken the one-to-one requirement from `BRL`; our sampler does ensure that each record in $B$ can be matched to at most one record in $A$, but allows for the possiblity that multiple records in $B$ match to the same record in $A$. We explore the ramifications of this distinction with the El Salvador case study.

@@

We calculate a Bayes estimate $\hat{\mathbf{Z}}$ for the linkage parameter $\mathbf{Z}$ by assigning different positive losses to different types of errors, and minimizing posterior expected loss. We adopt the loss function proposed in Sadinle 2017, in which $\hat{Z}_j \in \{1, \ldots, n_A + 1, R\}$, with $R$ representing the option to leave the matching undetermined by the model. Specifically, we minimize the quantity $L(\hat{\mathbf{Z}}, \mathbf{Z}) = \sum_{j=1}^{n_B} L(\hat{Z_j}, Z_j)$ where

$$L(\hat{Z_j}, Z_j)=\begin{cases} 
    0  & \text{if } Z_j = \hat{Z_j}; \\
    \theta_R,  & \text{if } \hat{Z_j} = R; \\
    \theta_{10},  & \text{if } Z_j \leq n_A,\hat{Z_j} = n_A + 1 ; \\
    \theta_{01},  & \text{if } Z_j = n_A + 1,\hat{Z_j} \leq n_A ; \\
    \theta_{11},  & \text{if } Z_j, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j} ; \\
\end{cases}$$
Here, $\theta_R$ is the loss from not making a decision on the linkage status, $\theta_{10}$ is the loss from a false non-match, $\theta_{01}$ is the loss from a false match, and $\theta_{11}$ is the loss from the special case of a false match in which the record has a true match other than the one estimated by the model. This loss function leads to closed form decision rules for minimizing posterior expected loss. In this paper, we adopt losses $\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2$, inducing the intuitive decision rule

$$\hat{Z}_j =\begin{cases} 
    i,  & \text{if } P(Z_j = i |\Gamma) > \frac{1}{2}; \\
    0,  & \text{otherwise} ; \\
\end{cases}$$

For a more indepth explanation of this function and the induced Bayes estimate, see Sadinle 2017.

Since our Gibbs procedure does not strictly enforce one-to-one matching, it is possible for the final Bayes estimate to link multiple records in $B$ to one record in $A$. The modeler can either report both such matches (with their respective posterior match probabilities), or resolve these conflicts by accepting only the match with highest posterior probability. Such a resolution procedure indeed is equivalent to minimizing posterior risk under the restriction of one-to-one matching, and as thus theoretically justified within the Bayesian framework.

@@

Broadly speaking, we increase our computational efficiency by recognizing that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector. Let $\mathcal{H}$ be the set of unique agreement patterns in the data, let $P = |\mathcal{H}|$ denote the total number of unique agreement patterns. Note that $P$ is bounded above by $\prod_{f=1}^F L_f$, and that this bound does not scale with $n_A$ or $n_B$. Prior to processing the data, we identify all $P$ patterns in $\mathcal{H}$ and enumerate them $h_1, \ldots, h_P$. When the $(i,j)$ pair exhibits the $p^{th}$ agreement pattern, we say $(i,j) \in h_p$. Wherever possible, we conduct calculations over these $P$ agreement patterns rather than the $n_A \times n_B$ record pairs.

@@

The classic Fellegi Sunter method represents the $\gamma_{ij}$ comparison vector as vector of length $F$, with each component $\gamma_{ij}^f$ taking on values in  $\{1, \ldots, L_f \}$. To ease computations, we instead represent the comparison as a concatenation of $F$ many binary indicator vectors of lengths $L_f$. For example, if $L_1 = L_2 = 2$ and $L_3 = 3$, then $\gamma_{ij} = (2, 1, 3)$ under the classical framework becomes $\gamma_{ij} = (0, 1, 1, 0, 0, 0, 1)$ under our framework. This is a bijective transformation that does not change the meaning of the data, but this representation will ease calculations and posterior updates.

In the classic Fellegi Sunter framework, $\Gamma$ is a $n_A n_B \times F$ matrix, with each row providing the comparison vector for a different $(i,j)$ pair. In contrast, we do not store these comparison vectors themselves, but instead only a hashed value corresponding to the agreement pattern of the $(i, j)$ pair. We store this information in a nested list $\tilde{\Gamma}$ where the $p^{th}$ component of the $j^{th}$ list contains a vector of records in $A$ that share agreement pattern $p$ with record $j \in B$. For each $p$, we also calculate $H_p = \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} \mathbf{1}_{(i,j) \in h_p}$, the total instances of agreement pattern $p$ throughout the data, and also for each $j$, we  calculate $H_{p_j} = \sum_{i=1}^{n_A} \mathbf{1}_{{(i,j) \in h_p}}$ the instances of agreement pattern $p$ among the comparison vectors between record $j \in B$ and each of the $n_A$ records in $A$.

@@

To address this, we propose a new strategy called *storage efficient indexing* (SEI). In standard indexing, the modeler decides a certain criteria that they expect all true matching pairs to satisfy, and a priori label any record pairs that do not meet that criteria as non-matching. For example, one might only consider pairs with a certain similarity score on a field deemed to be important (like first name), or only pairs with exact matching on a specified number of fields. While generally chosen to be be quite loose, establishing these criteria requires knowledge of the problem and invites room for human error. We propose a method of reducing the comparison space (and reducing storage requirements) without these drawbacks. Note that all records of the same agreement pattern have the same probability when sampling $Z_j$. Therefore we know that records belonging to an $h_p$ such that $H_{p_j}$ is large are very unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate, even without considering the form of the agreement pattern itself.

In SEI, rather than store all of these unlikely record labels, we choose to store only a small number $R$ of them. Posterior calculations still attribute the appropriate weight to all records through the summary statistics $H_p$, and $H_{p_j}$. Rather than storing $n_A \times n_B$ record labels, SEI allows us to store at most $n_B \times P \times R$ labels, regardless of how large $n_A$ is.

Laslty, for large data, we can partition the two datasets $A$ and $B$ into smaller blocks $\{A_m\}$ and $\{B_m\}$ for more manageable computations. On a single machine, we can read-in data sequentially, conduct hashing, compress information through SEI, and delete the original data from memory before continuing with the next chunk of data. With multiple cores or multiple machines, this can be done in parallel. Thus the combination of hashing, SEI, and partitioning allows us to conduct linkage tasks over much larger datasets.

@@

\underline{Updating $\mathbf{Z}$:} Although sampling $Z_j$ from a the full conditional provided earlier is conceptually straightforward, it becomes computational burdensome when $n_A$ is larger. The reader can confirm that sampling a value from a large set of unequal probabilities becomes difficult in most programming languages. To speed up computation, we break this sampling step into two simpler steps. First, we calculate the Fellegi Sunter weight $w_{p}$ associated with each unique pattern and sample the agreement pattern between $j$ and its potential match. Second, we sample the record label uniformly among records associated with that agreement pattern for that particular $j\in B$. More concretely, define $h(Z_j)$ to be the agreement pattern between $j$ and its potential match, and say $h(Z_j) = h_{P+1}$ when $Z_j = n_A + 1$. Then,

$$P\left(h\left(Z_j^{(s+1)}\right) = p | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}}\right) \propto
\begin{cases} 
    w_{p}\times H_{p_j}  & p \leq P; \\
     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &   p = P + 1 \\
\end{cases}$$
$$P\left(Z_j^{(s+1)} = i\bigg|\;\; h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
\frac{1}{H_{p_j}} & (i, j) \in h_p \\
0 & \text{otherwise} \\
\end{cases}$$

Lastly, we recognize that all posterior updates are governed by the agreement patterns of the record pairs rather than the record labels themselves. Thus we complete the entire Gibbs procedure first at the level of the $P$ agreement patterns with the first equation above. After, we can back-fill the records corresponding to the agreement patterns through the second equation. Sampling uniformly is computational simple even for large sets of candidate records, but this step can also be parallelized when working with large data.

@@

Comparing our method against `BRL` on these same simulated datasets, we find that our method only has weakened performance in the most extreme scenario of very high errors and very high overlap across files. We use the following measures of accuracy in our analysis: although using only recall and precision is common in the literature, we find it useful to have a single metric to gauge overall performance, and thus also include F measure.

@@

```{r fig.cap="Posterior means and credible intervals for accuracy metrics under the replication of simulation study from Sadinle 2017. For each level of overlap and each level of error, we have 100 paired sets of 500 records."}
knitr::include_graphics("../notes/figures/sadinle_sim_plot.png")
```

@@

To demonstrate speed, we generate comparison vectors from pre-specified distributions so that we can easily increase the size of the linkage problem. Distributions are meant to emulate the behavior of similarity scores across first name, last name, and day, month. For simplicity, we consider only exact matching, so a vector (1, 0) corresponds to agreeement and (0, 1) to disagreement.  We simulate these data for different values of $n_A$ and $n_B$, and compare the run-time of `fabl` against `BRL`. Note that the number of unique patterns $P$ is bounded above by $2^5 = 32$, a bound which is consistently attained in the larger simulations.

We see that at low data size, `BRL` outperforms, but that `fabl` is significantly faster at handling larger data. In particular, run-time for `BRL` seems to grow quadratically (or linearly with the size of both $A$ and $B$) while run-time for `fabl` seems to grow linearly (in the size of $B$).

\begin{table}[ht]
\centering
\begin{tabular}{rrr}
  \hline
 & m & u \\ 
  \hline
  fname & (0.95, 0.05) & (0.01, 0.99) \\ 
  lname & (0.95, 0.05) & (0.01, 0.99) \\ 
  day & (0.95, 0.05) & (0.03, 0.97) \\ 
  month & (0.95, 0.05) & (0.08, 0.92) \\ 
  year & (0.95, 0.05) & (0.07, 0.93) \\ 
   \hline
\end{tabular}
\end{table}

```{r fig.cap="Runtime for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_A$ and $n_B$"}
knitr::include_graphics("../notes/figures/sadinle_speed_plot2.png")
```

The above discussion suggests that for fixed $n_B$, computation time should remain mostly constant with growing $n_A$. Simulation study suggests that this is true. In the plot below, fixing $n_B = 500$, we see linear growth for the run-time under `BRL` as $n_A$ increases, with much more static run-time under `fabl`. The slight increases in run-time that we do see are due primarily to the hashing step, which again can be run in parallel for large data.

```{r fig.cap="Runtime for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_A$, and $n_B$ fixed at 500"}
knitr::include_graphics("../notes/figures/speed_plot_fixed_nB.png")
```
